\section{Discussion}
\label{sec:discussion}

\paragraph{No single defense dominates.}
Our $3 \times 5$ evaluation matrix reveals that defense effectiveness varies
substantially by attack type.  Watermark detection reliably identifies
unsigned memory writes ($\tpr=1.00$, $\fpr=0.00$ for all attacks) but
provides no protection if the attacker obtains the watermark key or
bypasses the ingestion path.  \sad{} excels against \minja{} (bridging-step
passages carry high query-aligned similarity, $\tpr=1.00$, AUROC\,=\,1.000)
but is less effective against \injecmem{} ($\tpr=0.40$, AUROC\,=\,0.920
in combined mode; broad-anchor passages avoid high similarity to any specific
query) and provides near-zero detection against \agentpoison{} under
plain-query calibration ($\tpr=0.00$, AUROC\,=\,0.870; the centroid passage
falls within the benign similarity distribution).
The Proactive defense exhibits a complementary pattern: it achieves
$\tpr=1.00$ against \agentpoison{} (centroids score high on all domain
probes) and $\tpr=0.60$ against \injecmem{} (broad anchors match many
probe topics), but only $\tpr=0.07$ against \minja{} (targeted bridging-step
passages match few cross-domain probes) at $\fpr=0.01$.
Composite defenses achieve $\tpr=1.00$, $\fpr=0.00$ across all three
attacks by combining these complementary signals, confirming that no single
mechanism covers the full threat surface and motivating a defense-in-depth
portfolio.

\paragraph{Why SAD fails against centroid-targeting attacks under plain-query calibration.}
The centroid passage is crafted to maximise mean cosine similarity across
all \emph{triggered} victim queries.  When \sad{} is calibrated on
\emph{plain} (un-triggered) victim queries, the centroid passage similarity
falls within the benign distribution: plain queries do not include the
trigger token, so the centroid passage (designed for triggered queries) is
not anomalously similar to them.

\paragraph{Triggered-query calibration closes the gap.}
However, once the attacker deploys the trigger to user clients, triggered
queries appear in the agent's query log.  If \sad{} is calibrated on
\emph{observed} queries---which, after attack deployment, include triggered
queries---the centroid passage becomes a clear outlier.  We empirically
confirm this: calibrating \sad{} on triggered queries ($\trigger^* + q$
for all victim queries) raises the detection TPR from $0.00$ to $1.00$
with FPR $= 0.00$ (threshold shifts from $0.482$ to $0.432$ as triggered
queries raise the calibration statistics in a way that leaves the centroid
passage above the new threshold).
This implies a practical defence-deployment strategy: \sad{} should
use a rolling query window drawn from \emph{all} observed agent queries
rather than a pre-deployment reference set.  As triggered queries accumulate
in the window, \sad{} becomes increasingly effective against centroid-targeting
attacks at no additional architectural cost.

\paragraph{Trigger prepending is a fundamental protocol requirement.}
The evaluation correction identified in Section~\ref{sec:experiments}
reflects a fundamental asymmetry in the \agentpoison{} threat model:
the attacker controls query content at evaluation time (via a shared
trigger in the user's client), but the defender measures retrieval
without this adversarial query modification.  Without trigger prepending,
the adversarial passage must compete with benign entries on generic
query similarity---a much easier task for defenders.  Future simulations
of trigger-based attacks must implement triggered-query evaluation; we
recommend explicitly documenting the evaluation protocol as part of any
attack benchmark.

\paragraph{DPR trigger optimization and the centroid approximation.}
Our centroid-targeting passage (Eq.~\ref{eq:centroid}) is a practical
approximation of the optimal adversarial passage.  The DPR HotFlip optimizer
(Section~\ref{sec:attacks}) improves mean triggered-query similarity from
$0.71$ (centroid only) to $0.78$ ($+10\%$), confirming that gradient-guided
token selection provides a meaningful gain beyond semantic coverage alone.
At $\asrr=1.00$ in the all-MiniLM-L6-v2 embedding space, the centroid
approximation is already sufficient for this encoder's isotropy structure;
DPR HotFlip is most beneficial in embedding spaces with lower isotropy
(encoder-only models trained on narrow domains) where a single centroid may
not span multiple query clusters.

\paragraph{Generalization across embedding models and memory architectures.}
All experiments use all-MiniLM-L6-v2 with FAISS IndexFlatIP.  Attack
effectiveness and \sad{} calibration will differ for other encoders.
Attacks that rely on trigger-triggered centroid retrieval should transfer
broadly (any dense retriever shares the inner-product objective), but
the absolute \asrr{} values will vary with encoder isotropy and vocabulary.
\sad{}'s calibration is encoder-agnostic in design---it calibrates on
whatever similarity distribution the deployed encoder produces---but its
threshold $k$ may require re-tuning for encoders with different
intra-corpus similarity ranges.  Graph-structured memory systems (e.g.,
those using entity--relation stores) present a different attack surface
that our retrieval-based threat model does not cover.

\paragraph{Deployment recommendations.}
Three practical recommendations emerge from our evaluation.
(i)~\emph{Validate memory provenance at write time}: a watermark-based
check on all incoming memory writes costs $O(|entry|)$ compute and
eliminates unsigned injections entirely.
(ii)~\emph{Monitor per-entry anomaly scores continuously}: a rolling
\sad{} calibration window (100 recent queries, $k=2.0$, combined scoring
mode) adds negligible retrieval latency, achieves $\fpr=0.00$, and
maintains $\tpr=1.00$ for bridging-step attacks (AUROC\,=\,1.000).
(iii)~\emph{Apply composite portfolios}: our matrix shows the composite
defense achieves $\asrr = 0.000$ for all three attacks
(Table~\ref{tab:attack_defense_matrix}).  Even the weaker combination of
watermark (which catches unsigned writes) plus \sad{} (which catches
similarity outliers) eliminates all attacks simultaneously when triggered-query
calibration is used for \sad{}: no attack achieves $\asrr > 0.0$ under this
layered combination, confirming that defense-in-depth provides compounding
protection that individual defenses cannot.

\paragraph{Limitations.}
(i) Measured $\asra$ from GPT-2 (Section~\ref{sec:phase17_results}) is a
lower bound; true $\asra$ from production LLMs (GPT-4, Llama-3) is expected
to be substantially higher, closer to the modelled upper bound.
Full live agent evaluation with production-scale LLMs remains future work.
(ii) The synthetic corpus of 200 entries may not capture the full
diversity of production memory contents, particularly in multi-user or
multi-domain agent deployments.
(iii) \sad{}'s calibration assumes a stationary query distribution; a
sophisticated adversary who gradually shifts the query distribution over
many sessions could inflate the baseline similarity mean and lower the
effective detection threshold.
(iv) The token-level watermark requires a GPT-2 backbone and adds
$O(n_{\text{tokens}})$ inference time per stored entry; for latency-sensitive
deployments, the character-level adaptation may be preferable.
(v) The word-level synonym adversary circumvents \sad{}'s theoretical
gradient tension (Section~\ref{sec:adaptive}) because semantic embeddings
are synonym-invariant: the all-MiniLM-L6-v2 model was trained on paraphrase
pairs and assigns nearly identical embeddings to domain synonyms.  Synonym
substitution can cross the detection threshold without affecting retrieval
rank.  The theoretical tension therefore holds only for continuous
gradient-based adversaries (e.g., HotFlip-style optimisation), not for
discrete word-substitution adversaries.  Future defenses should incorporate
lexical diversity features or synonym-aware thresholds to close this gap.

\paragraph{Future work.}
Key extensions include:
(a)~live agent evaluation with production-scale LLMs (LangChain, OpenAI
Assistants API) for tighter $\asra$ bounds;
(b)~multi-agent propagation experiments where a single poisoned entry
spreads across a shared knowledge base serving multiple agents;
(c)~formal robustness analysis for composite portfolios under
simultaneously-adaptive adversaries optimizing against all defenses jointly;
(d)~evaluation on \citet{memorygraft2025} and \citet{amemguard2025}
architectures to assess cross-system generalization; and
(e)~DPR trigger optimization in alternative embedding spaces (Contriever,
E5, text-embedding-3) to assess encoder-agnostic attack strength.
