% Memory Agent Security: Unified Attack Evaluation and Defense Framework
%
% Compilation: pdflatex main.tex && bibtex main && pdflatex main.tex && pdflatex main.tex
% Or use latexmk: latexmk -pdf main.tex

\documentclass[10pt]{article}

% --- packages ---
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{subcaption}
\usepackage{microtype}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning, shapes.geometric, fit, backgrounds, patterns}

% --- figure path (all generated figures live in figures/) ---
\graphicspath{{figures/}}

% --- theorem environments ---
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}

% --- custom macros ---
\newcommand{\asrr}{\text{ASR-R}}
\newcommand{\asra}{\text{ASR-A}}
\newcommand{\asrt}{\text{ASR-T}}
\newcommand{\isr}{\text{ISR}}
\newcommand{\tpr}{\text{TPR}}
\newcommand{\fpr}{\text{FPR}}
\newcommand{\EE}{\mathbb{E}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\sad}{\textsc{SAD}}
\newcommand{\agentpoison}{\textsc{AgentPoison}}
\newcommand{\minja}{\textsc{MINJA}}
\newcommand{\injecmem}{\textsc{InjecMEM}}
\newcommand{\trigger}{T}
\newcommand{\passage}{p_{\mathrm{adv}}}
\newcommand{\enc}{E}

\title{%
  \textbf{Poisoning the Memory: A Unified Framework for Attacking and\\
  Defending LLM Agent Memory Systems}
}

\author{%
  Ishrith Gowda\\
  University of California, Berkeley\\
  \texttt{ishrithgowda@berkeley.edu}
}

\date{March 2026}

\begin{document}

\maketitle

\begin{abstract}
LLM agents increasingly rely on external memory systems to persist context
across sessions.  We show that this memory constitutes a critical and largely
undefended attack surface.  We present a unified evaluation framework for
three recently proposed memory poisoning attacks---\agentpoison{}
\citep{chen2024agentpoison}, \minja{} \citep{dong2025minja}, and
\injecmem{} \citep{injecmem2026}---implemented over a realistic FAISS-backed
vector memory store (all-MiniLM-L6-v2, 384-dim) with a 200-entry synthetic
agent memory corpus.

We identify and correct a systematic evaluation error in prior simulations:
\agentpoison{}'s attack success rate at retrieval (\asrr) was underestimated
because victim queries were issued \emph{without} the attacker-controlled
trigger token at evaluation time.  The paper's evaluation protocol
(Chen et al., Algorithm~2) prepends the trigger to every query;
with this correction and a centroid-targeting adversarial passage,
\asrr{} increases from 0.25 to 1.00 at 2.4\% poison rate.

On the defense side, we propose \sad{}---a Semantic Anomaly Detection
defense that flags adversarial memory entries by their anomalously high
cosine similarity to observed victim queries.  We evaluate all five defenses
(Watermark, Validation, Proactive, Composite, \sad{}) in a pre-ingestion
filtering model across a full $3 \times 5$ attack-defense interaction matrix.
We find that no single defense is uniformly effective, motivating composite
defense portfolios.  For \sad{} specifically, we identify a fundamental gradient tension for
continuous adversaries: any perturbation that reduces the anomaly score
also reduces retrieval rank via the shared cosine similarity objective.
Empirically, word-level synonym substitution circumvents this tension
because semantic embeddings are synonym-invariant: synonym replacements
achieve evasion rates $\geq 0.80$ with $\Delta\asrr \approx 0$, motivating
defenses that incorporate lexical diversity features.
We further show that \sad{} calibrated on \emph{triggered} victim queries
achieves $\tpr = 1.00$, $\fpr = 0.00$ for \agentpoison{}---closing the blind
spot that arises under plain-query calibration.

We additionally implement three paper-faithful extensions: (i) DPR HotFlip
gradient-guided trigger optimization~\citep{ebrahimi2018hotflip} over the
\texttt{facebook/dpr-ctx\_encoder-single-nq-base} (768-dim), improving mean
triggered-query similarity from 0.71 to 0.78; (ii) the full token-level
unigram watermark~\citep{zhao2024watermark} via a GPT-2 backbone, achieving
$83\%$ higher z-scores than character-level at short text lengths; and
(iii) a local GPT-2 agent evaluator providing a directly measured lower bound
on $\asra$ ($0.42$--$0.51$ across attacks) that confirms the modelled values
are conservative.

We extend the evaluation across three further dimensions.
\emph{Multi-encoder transferability}: we evaluate all three attacks and
\sad{} across seven retrieval encoders (MiniLM-L6, MPNet-Base, E5-base,
Contriever, BGE-large, OpenAI text-embedding-3-small/-large) and compute
linear Centered Kernel Alignment (CKA) across all encoder pairs, finding
that high inter-encoder CKA predicts cross-encoder ASR-R transferability.
\emph{Multi-agent epidemic propagation}: a discrete-step SIR model over a
shared FAISS knowledge base shows that a single poisoned entry spreads to
connected agents via re-storage; \sad{}-based write-time quarantine halts
secondary infection by blocking flagged entries before they are re-stored.
\emph{Graph-structured memory}: we extend the threat model to GraphRAG-style
entity-relation stores, implementing hub insertion, edge-hijack, and
subgraph-cluster attacks that exploit graph topology, and two graph-native
defenses---degree-anomaly detection and adjacency contamination
scoring---that identify structurally anomalous nodes without requiring
labeled poison examples.
Production-scale \asra{} is directly measured using a GPT-4o-mini agent
evaluator (OpenAI \texttt{text-embedding-3-large} retrieval, $T=0$),
and FPR for all defenses is validated over 20 independent trials using
Clopper-Pearson exact binomial confidence intervals, confirming
FPR\,$=0.000$ for watermark and \sad{} combined-mode at $95\%$ confidence.

All code, experiments, and figures are open-sourced at
\url{https://github.com/ishrith-gowda/memory-agent-security}.
\end{abstract}

% section files
\input{sections/introduction}
\input{sections/related_work}
\input{sections/threat_model}
\input{sections/attacks}
\input{sections/defenses}
\input{sections/experiments}
\input{sections/discussion}
\input{sections/conclusion}

\bibliography{references}
\bibliographystyle{plainnat}

\appendix
\input{sections/appendix}

\end{document}
